---
title: "BigHornSheep"
author: "Julie Norman"
date: '2022-03-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Upload

### Libraries

```{r}
library(tidyr)
library(tidyverse)
library(stringr)
library(mapview)
library(sf)
library(sp)

library(rgeos)
```

copy coordinates from pdf and save as .txt file. Delete all line breaks before uploading using find/replace function

```{r}
txt1 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit1.txt", header = FALSE, sep = ";")) # transposed to long
txt1_df <- as.data.frame(txt1) # convert to df
txt1_df$Unit <- "1" # add area unit


txt2 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit2.txt", header = FALSE, sep = ";"))
txt2_df <- as.data.frame(txt2) 
txt2_df$Unit <- "2"

txt3 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit3.txt", header = FALSE, sep = ";"))
txt3_df <- as.data.frame(txt3) 
txt3_df$Unit <- "3"

txt4 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit4.txt", header = FALSE, sep = ";"))
txt4_df <- as.data.frame(txt4) 
txt4_df$Unit <- "4"

txt5 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit5.txt", header = FALSE, sep = ";"))
txt5_df <- as.data.frame(txt5) 
txt5_df$Unit <- "5"

txt6 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit6.txt", header = FALSE, sep = ";"))
txt6_df <- as.data.frame(txt6) 
txt6_df$Unit <- "6"

txt7 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit7.txt", header = FALSE, sep = ";"))
txt7_df <- as.data.frame(txt7) 
txt7_df$Unit <- "7"

txt8 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit8.txt", header = FALSE, sep = ";"))
txt8_df <- as.data.frame(txt8) 
txt8_df$Unit <- "8"

txt9 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit9.txt", header = FALSE, sep = ";"))
txt9_df <- as.data.frame(txt9) 
txt9_df$Unit <- "9"

txt10 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit10.txt", header = FALSE, sep = ";"))
txt10_df <- as.data.frame(txt10) 
txt10_df$Unit <- "10"

txt11 <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit11.txt", header = FALSE, sep = ";"))
txt11_df <- as.data.frame(txt11) 
txt11_df$Unit <- "11"

```

Combine into one dataset
```{r}
cord_df <- rbind(txt1_df, txt2_df) %>% 
  rbind(txt3_df) %>% 
  rbind(txt4_df) %>% 
  rbind(txt5_df) %>% 
  rbind(txt6_df) %>% 
  rbind(txt7_df) %>% 
  rbind(txt8_df) %>% 
  rbind(txt9_df) %>% 
  rbind(txt10_df) %>% 
  rbind(txt11_df) 
```


Reformat data frame

```{r}

rownames(cord_df) <- NULL # delete row names

colnames(cord_df) <- c("Coordinates", "Unit") # add name columns

cord_df <- cord_df %>% na.omit # delete empty rows


cord_df
```
## Create Shape Files

separate coordinates into latitude and longitude columns 

```{r}
cord_df <- separate(cord_df, "Coordinates", c("lat", "lon"), ",")

# convert from factor to numeric
cord_df$lat <- as.numeric(cord_df$lat)
cord_df$lon <- as.numeric(cord_df$lon)
cord_df$Unit <- as.numeric(cord_df$Unit)

cord_df
```

Crs gathered based on GeodeticaDA and utmZone
https://spatialreference.org/ref/epsg/nad83-utm-zone-11n/ 




## Create Spacial file 

```{r}

# make a list
cord_list <- split(cord_df, cord_df$Unit)

# extract lon-lats in the list, not the names
cord_list <- lapply(cord_list, function(x) { x["Unit"] <- NULL; x })


# create SpatialPolygons Object, convert coords to polygon
ps <- lapply(cord_list, Polygon)


# add Unit variable
p1 <- lapply(seq_along(ps), function(i) Polygons(list(ps[[i]]), 
                                            ID = names(cord_list)[i]  ))

# create spatial polygons
bhs_sp <- SpatialPolygons(p1, proj4string = CRS("+proj=utm +zone=11 +ellps=GRS80 +datum=NAD83 +units=m +no_defs") )


# create spatial polygon df
bhs_spdf <- SpatialPolygonsDataFrame(bhs_sp, 
 data.frame(Unit = unique(cord_df$Unit), 
            row.names = unique(cord_df$Unit)))

# convert to shapefile
bhs_sf <- st_as_sf(bhs_spdf, crs = "+proj=utm +zone=11 +ellps=GRS80 +datum=NAD83 +units=m +no_defs ")
```

## Add removed areas

### Upload removed areas data

```{r}
txt1A <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit1A.txt", header = FALSE, sep = ";")) # transposed to long
txt1A_df <- as.data.frame(txt1A) # create dataframe
txt1A_df$Unit <- "1" # add Unit variable

txt4A <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit4A.txt", header = FALSE, sep = ";"))
txt4A_df <- as.data.frame(txt4A) 
txt4A_df$Unit <- "4A"

txt4B <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit4B.txt", header = FALSE, sep = ";"))
txt4B_df <- as.data.frame(txt4B) 
txt4B_df$Unit <- "4B"

txt4C <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit4C.txt", header = FALSE, sep = ";"))
txt4C_df <- as.data.frame(txt4C) 
txt4C_df$Unit <- "4C"

```

Combine into one dataset
```{r}
crmv_df <- rbind(txt1A_df, txt4A_df) %>% 
  rbind(txt4B_df) %>% 
  rbind(txt4C_df) 
```

adjust row and column names
```{r}
rownames(crmv_df) <- NULL # delete row names

colnames(crmv_df) <- c("Coordinates", "Unit") # name columns

crmv_df <- crmv_df %>% na.omit # delete empty rows

# separate coordinates into two columns
crmv_df <- separate(crmv_df, "Coordinates", c("lat", "lon"), ",") 

# convert to numeric
crmv_df$lat <- as.numeric(crmv_df$lat)
crmv_df$lon <- as.numeric(crmv_df$lon)

```

### Creating removed areas

```{r}

# make a list
crmv_list <- split(crmv_df, crmv_df$Unit)

# extract lon-lats in the list, not the names
crmv_list <- lapply(crmv_list, function(x) { x["Unit"] <- NULL; x })

# create SpatialPolygons Object, convert coords to polygon
rm_ps <- lapply(crmv_list, Polygon)


# add Unit variable
rm_p1 <- lapply(seq_along(rm_ps), function(i) Polygons(list(rm_ps[[i]]), 
                                            ID = names(crmv_list)[i]  ))

# convert to spatial polygons
rmbhs_sp <- SpatialPolygons(rm_p1, proj4string = CRS("+proj=utm +zone=11 +ellps=GRS80 +datum=NAD83 +units=m +no_defs") )


# create spatial polygons dataframe
rmbhs_spdf <- SpatialPolygonsDataFrame(rmbhs_sp, 
 data.frame(Unit = unique(crmv_df$Unit), 
            row.names = unique(crmv_df$Unit)))

# convert to spatial file
rmbhs_sf <- st_as_sf(rmbhs_spdf, crs = "+proj=utm +zone=11 +ellps=GRS80 +datum=NAD83 +units=m +no_defs ")
```

### Subtract Polygons

```{r}
Unit1 <- filter(bhs_sf, Unit == 1) # Unit 1 polygon
Unit1A <- filter(rmbhs_sf, Unit == 1) # Unit 1 removed section polygon

# subtract polygons, grab relavent columns
Unit1_comp <- st_difference(Unit1, Unit1A) %>% 
  select(Unit, geometry)

```

```{r}
# grab individual polygons
Unit4 <- filter(bhs_sf, Unit == 4)
Unit4A <- filter(rmbhs_sf, Unit == "4A")
Unit4B <- filter(rmbhs_sf, Unit == "4B")
Unit4C <- filter(rmbhs_sf, Unit == "4C")

# subtract removed sections, grab columns
Unit4_comp <- st_difference(Unit4, Unit4A) %>% 
  st_difference(Unit4B) %>% 
  st_difference(Unit4C) %>% 
  select(Unit, geometry)

```



### Add new polygons 

```{r}
# remove Unit 1 and 4 main polygons
bhs2_sf <- filter(bhs_sf, Unit != 1 & Unit != 4)

# add Units 1 and 4 with removed sections
bhsFinal_sf <- rbind(bhs2_sf, Unit1_comp) %>% 
  rbind(Unit4_comp)
  
```

```{r}
`Bighorn Sheep Habitat` <- bhsFinal_sf # rename for legend title

mapview(`Bighorn Sheep Habitat`)
```