---
title: "MD1"
author: "Julie Norman"
date: '2022-03-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Upload Coordinates

### Libraries

```{r}
library(tidyr)
library(tidyverse)
library(stringr)
library(mapview)
library(sf)

library(sp)
```


copy coordinates from pdf and save as .txt file. Delete all line breaks beforehand using find/replace function

```{r}
txt <- t(
  read.csv("/Users/julienorman/Documents/Economics/CGO\ Projects/CriticalHabitatMap/BHS_Unit1.txt", header = FALSE, sep = ";"))

txt_df <- as.data.frame(txt) # convert to df

#flip orientation with and add column names
rownames(txt_df) <- NULL 

colnames(txt_df) <- "Coordinates"

txt_df <- txt_df %>% na.omit # delete empty rows

str(txt_df)
```


separate coordinates into latitude and longitude columns 

```{r}
cord_df <- separate(txt_df, "Coordinates", c("lat", "lon"), ",")

cord_df
```
### Add other dataset information

geodeticDA and utmZone values found in pdf


```{r}
cord_df$geodeticDA <- "NAD83"
cord_df$utmZone <- "11N"
cord_df$Unit <- "1"

cord_df
```

Crs gathered based on GeodeticaDA and utmZone
https://spatialreference.org/ref/epsg/nad83-utm-zone-11n/ 

### Create sf file and plot

```{r}
plot_bhs1 <- st_as_sf(cord_df, 
                        coords = c("lat", "lon"),
                         crs = "+proj=utm +zone=11 +ellps=GRS80 +datum=NAD83 +units=m +no_defs ")

utm11nCRS <- st_crs(plot_bhs1)

```


```{r}
mapview(plot_bhs1, map.types = "Esri.NatGeoWorldMap") 
```


# Create Polygon

```{r}
polygon <- plot_bhs1 %>% 
  dplyr::group_by(Unit) %>% 
  dplyr::summarise() %>%
  sf::st_cast("POLYGON")
polygon
```

```{r}
mapview(polygon, map.types = "Esri.NatGeoWorldMap") 
```



```{r}
polys <- st_sf(
  aggregate(
    plot_bhs1$geometry,
    list(plot_bhs1$Unit),
    function(g){
       st_cast(st_combine(g),"POLYGON")
    }
   ))
```

```{r}
mapview(polygon, map.types = "Esri.NatGeoWorldMap") 
```



# testing

```{r}

cord_df <- separate(txt_df, "Coordinates", c("lat", "lon"), ",")

df <- cord_df

df$lat <- as.numeric(df$lat)
df$lon <- as.numeric(df$lon)

sp_df <- SpatialPoints(df)


plot(sp_df)

```

```{r}
p = Polygon(df)
ps = Polygons(list(p),1)
sps = SpatialPolygons(list(ps))
plot(sps)
```

```{r}
proj4string(sps) = CRS("+proj=utm +zone=11 +ellps=GRS80 +datum=NAD83 +units=m +no_defs")

data = data.frame(Unit=1)
spdf = SpatialPolygonsDataFrame(sps,data)
spplot(spdf)
```

```{r}
mapview(spdf, map.types = "Esri.NatGeoWorldMap") 
```

```{r}
test <- st_as_sf(spdf)
```

